---
title: "Scrape titanic list"
---

The site [Titanic Facts](https://titanicfacts.net/titanic-passenger-list/) has a full list of passengers and a list of just the survivors and their lifeboat.

My aim is to get a single list that I can use in class.

## Setup

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(rvest)
```

## Scrape the dead

Get the web page

```{r}
page_raw <- read_html("https://titanicfacts.net/titanic-passenger-list/")
```

Find the tables on that page

```{r}
tables_list <- page_raw %>% html_table()
```

The tables have bad column names. I fix that, add what class passenger they were based on the table they came from.

```{r}
c1_raw <- tables_list[[1]] %>% row_to_names(1) %>% mutate(class = "1st") %>% clean_names()
c2_raw <- tables_list[[2]] %>% row_to_names(1) %>% mutate(class = "2nd") %>% clean_names()
c3_raw <- tables_list[[3]] %>% row_to_names(1) %>% mutate(class = "3rd") %>% clean_names()

pass_combo <- bind_rows(c1_raw, c2_raw, c3_raw)
```

Here I adjust the survivor column to remove cross for V for victim. Also renames.

```{r}
passengers <- pass_combo %>%
  mutate(survivor_victim = if_else(
    survivor_s_or_victim == "S", "S", "V")
  ) %>%
  select(-survivor_s_or_victim)

passengers %>% glimpse()
```

## Survivors

The survivors list are the same folks as on the passenger list if they survived, with the addition of what lifeboat they were in.

Get the page

```{r}
surv_page_raw <- read_html("https://titanicfacts.net/titanic-survivors-list/")
```

Get the tables

```{r}
surv_tables <- html_table(surv_page_raw)
```

While it is really one table, they split it into three I guess so they can add a "back to top" link in between them. I just mush them all together.

I also have to fix the column names and then remove the extra columns names from the 2nd and 3rd tables.

```{r}
surv_raw <- surv_tables %>% list_rbind() %>% row_to_names(1) %>% clean_names()

survivors <- surv_raw %>% filter(class != "Class")

survivors %>% glimpse()
```

What is lifeboat?

```{r}
survivors %>% count(lifeboat_number)
```

## Combine these files

I'm combining the survivor list, which has the lifeboat, with the passenger list. I have to first drop the survivors from the passengers list then bind them back from the survivors list.

```{r}
titanic <- passengers %>% filter(survivor_victim == "S") %>% 
  bind_rows(survivors)

titanic %>% slice_sample(n = 20)
```

## Exporting files

I export the individual files, and then the combined ones as both rds and csv.

```{r}
passengers %>% write_rds("data/tf_passengers.rds")
survivors %>% write_rds("data/tf_survivors.rds")
titanic %>% write_rds("data/tf_titanic.rds")
titanic %>% write_csv("data/tf_titanic.csv")
```


